---
title: "Workshop Title"
output: html_notebook
---

## Learning Objectives

Welcome to the D-Lab GitHub Copilot Workshop. After this workshop, you will be able to:

1.  Set-up and navigate Visual Studio Code
   
2.  Take advantage of the main functionalities of GitHub Copilot for coding
   
3.  Understand some of the strengths and weaknesses of AI coding assistants
  
Throughout this workshop series, we will use the following icons:

🔔 **Question**: A quick question to help you understand what's going on.

🥊 **Challenge**: Interactive exercise. We'll go through these in the workshop!

⚠️ **Warning**: Heads-up about tricky stuff or common mistakes.

💡 **Tip**: How to do something a bit more efficiently or effectively.

📝 **Poll**: A zoom poll to help you learn.

🎬 **Demo**: Showing off something more advanced so you know what you can use R for in the future

## 1. Getting Comfortable with Visual Studio Code

Visual Studio Code (VS Code) is a free code editor. When you start VS Code, you will be greeted by a Welcome screen that prompts you to start by opening a file or folder. Click `Open Folder` and open the GitHub-Copilot workshop folder that you downloaded from GitHub. 

Now that you have opened a folder, you will see an `Explorer` panel appear with all of the files within the folder you just opened. Click on the `workshop_R.Rmd` file to open it. You can open and close the `Explorer` panel by clicking on the icon that looks like two pieces of paper in the left-hand sidebar.

💡 **Tip**: When we are working in a markdown file like this one, you may notice the text sometimes goes off the screen if the line is very long. We can change our VS Code settings to make it so that the text in our markdowns is automatically wrapped. To do this: 
1. Press `Ctrl + ,` to open the settings (You can also click on the cog icon in the bottom left corner of the window and then click `Settings`).
2. Search for word "wrap" in the search bar.
3. In the dropdown menu for `Editor: Word Wrap`, select on.

### Visual Studio Code Extensions
**Extensions** are add-ons that you can install to add new features to Visual Studio Code. You can find extensions by clicking on the square icon with four squares in the left-hand sidebar to open the `Extensions` panel. We will be using the following extensions in this workshop:
1. R - This extension provides support for the R programming language. You must have R already installed on your computer to use this extension.
2. GitHub Copilot - This extension is what gives us access to Copilot in VS Code

### 🥊 Challenge 1: 
In the search bar, type R. Click on the R extension and then click the green `Install` button. Repeat for the GitHub Copilot extension.

### Running code in Visual Studio Code
Now that we have all the extensions we need, we can start running R code! To run a line of code, click on the line you want to run and press `Ctrl + Enter`. To run multiple lines of code, highlight the lines you want to run and press `Ctrl + Enter`. If you are working in an R Markdown file, you can run a code chunk by clicking on the `Run Chunk` button that appears at the top of the chunk.

### 🥊 Challenge 2: 
Run the R code below and see what happens.
```{r}
print("Yippee!")
```

When you run this code you will notice that a new window appears below: the `Terminal`. This is where the output of your code will appear. You can also run code in the terminal by typing directly into it and pressing `Enter`. If you have used RStudio before, the terminal is similar to the console in RStudio.

Now that we know how to run R, let's install and load the necessary packages for this workshop. First, let's install the `tidyverse` package. You can install it by running the code below. We will use this today for manipulating our data and for making plots.

```{r}
# If you haven't installed tidyverse before, uncomment this line 
# You may be prompted to pick a CRAN mirror: just choose the first one (0-Cloud [https])
#install.packages("tidyverse")

# If you have previously installed tidyverse, load it into the current R session
library(tidyverse)
```

We will also need the `httpgd` package to display plots interactovely in VS Code. You can install it by running the code below.
```{r}
# if you haven't installed httpgd before, uncomment this line
#install.packages("httpgd")
```
💡 **Tip**: Unlike tidyverse or other packages you may actively use when coding, we do not need to load the `httpgd` package (i.e., we don't need to run `library("httpgd")`). `httpgd` just provides the infrastructure for showing plots in VS Code.

Once we have installed the `httpgd` package, you need to set `httpgd` as the default plot viewer. You can do this by: 
1. Press `Ctrl + ,` to open the settings 
2. Search for `r.plot.useHttpgd` in the search bar
3. Click on the tick box to set it to true

### 🥊 Challenge 3:
Check that both tidyverse and httpgd are installed and working by running the code below:
```{r}
# Create a tiny data.frame
tiny_df <- data.frame(x = c(1, 2, 3), y = c(4, 5, 6))

# Plot the data
ggplot(tiny_df, aes(x = x, y = y)) +
  geom_point()
```

## 2. Introduction to GitHub Copilot

Now that we have set up Visual Studio Code, we can start using GitHub Copilot!

For our examples, we will be using data from [Gapminder](https://www.gapminder.org/), an educational non-proﬁt. It includes data for 142 countries, with values for life expectancy, GDP per capita, and population, every five years, from 1952 to 2007.

```{r}
# read in data
gap <- read.csv(file = "data/gapminder.csv")
head(gap)
```

⚠️ **Warning**: By default, VS Code will set the working directory (i.e., where we "are" in your computer) to that of the folder you opened to start working in VS Code (in this case, the GitHub-Copilot folder). This is different than RStudio where the working directory is set by default to the location of your file or R Project. You can check where your working directory is by running `getwd()`. Try it out below:
```{r}
getwd()
```

### GitHub Copilot Chat

GitHub Copilot chat allows you to directly ask questions to GitHub Copilot. To open the chat, cick on the icon that looks like two text bubbles on the left-hand sidebar. You can ask Copilot to generate code, explain code, or provide examples. You can also start a chat by pressing `Ctrl + Shift + I`.

### 🥊 Challenge 4:
Open the chat, highlight the code above, and ask Copilot to explain it. Then ask Copilot to give you code to generate a summary of the data in the `gap` data frame by plotting histograms of "lifeExp" and "gdpPercap".

From Challenge 4 you should have gotten code that looks something like this:
```{r}
# Assuming 'gap' is your data frame
library(ggplot2)

# Histogram for lifeExp
ggplot(gap, aes(x = lifeExp)) +
  geom_histogram(binwidth = 5, fill = "blue", color = "black") +
  labs(x = "Life Expectancy", title = "Histogram of Life Expectancy")

# Histogram for gdpPercap
ggplot(gap, aes(x = gdpPercap)) +
  geom_histogram(binwidth = 1000, fill = "blue", color = "black") +
  labs(x = "GDP per Capita", title = "Histogram of GDP per Capita")
```

Note a few things:

1. It loads `ggplot2` because it knows we need it for plotting - however, this is unnecessary because we have loaded the `tidyverse` package which includes `ggplot2`. However, GitHub Copilot doesn't know that. 

2. It sets the binwidth for the histograms to 5 and 1000 - These are arbitrary values that Copilot chose and likely you as the user will want to change them. 

3. It provides guesses for the x axis title based on the name of the column in the data frame - here it has done a pretty good job! 

This is a good example of what Copilot can do for you, but also where you as the user need to be aware of what it is doing and make changes as necessary.

### In-line Chat
You can start an in-line chat with copilot by pressing `Ctrl + I`. This will allow you to ask Copilot for help with the code you are currently writing. 

When you activate in-line chat while highlighting code, Copilot will edit the code in the highlighted area.
### 🥊 Challenge 5: 
Highlight this ggplot code and ask it to change the binwidth to 500, change the fill color to your favorite color (e.g., "purple"), and change the background of the plot to white.
```{r}
ggplot(gap, aes(x = gdpPercap)) +
  geom_histogram(binwidth = 1000, fill = "blue", color = "black") +
  labs(x = "GDP per Capita", title = "Histogram of GDP per Capita")
```
⚠️ **Warning**: Sometimes you may notice that Copilot doesn't do everything you told it to. This is because Copilot is still learning and may not always understand what you are asking for. You can always ask it again or make the changes yourself.

### 🥊 Challenge 6: 
You can also use in-line chat in a blank line to generate new code. Start in-line chat in the empty chunk below and ask Copilot to generate code to calculate the mean of the "pop" column in the `gap` data frame grouped by the "country" column.
```{r}
# Start in-line chat here
```

### In-line suggestions
You may have noticed throughout this workshop that Copilot provides suggestions as you type in a lighter gray text. These suggestions can be helpful for completing code quickly. You can accept a suggestion by pressing `Tab` or `Enter`. 

Suggestions are automatically triggered by the code your write, based on context from the code you have already written. 

### 🥊 Challenge 6: 
Start typing `unique_countries <- gap` and see what suggestions Copilot gives you. Accept the suggestion that completes the code to create a vector of unique countries in the `gap` data frame.
```{r}
# Start typing here
```

⚠️ **Warning**: Copilot sometimes provides suggestions that are not what you want. You can always ignore the suggestion and continue typing. Copilot can change over time and give different suggestions based on even minor differences in context clues. You can test this out by deleting the line of code you just wrote and typing it again to see if you get the same suggestion.

### 🥊 Challenge 7:
You can also trigger suggestions by providing comments in your code. These commments allow you to give more detailed information and context for the in-line suggestions. Start typing `# Calculate average gdpPercap and lifeExp grouped by continent for the year 2007` and see what suggestions Copilot gives you. 

💡 **Tip**: You may have to start writing `gap %>%` to trigger the suggestions.
```{r}
# Start typing here
```

### DEMO idea: use copilot to generate a function

## Key Points

\<Summary of what was learned in today's workshop.\>

-   \<key point 1\>

-   \<key point 2\>

-   \<key point 3\>

-   max 5
